---
title: "Lab 5 - Penguin size comparisons"
author: "Jacey Van Wert"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(here)
library(effsize) #for cohen.d
library(broom)
library(car)
library(ggbeeswarm)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap using penguins dataset (preinstalled with {palmerpenguins})

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap") 


# Histograms
ggplot(chinstraps, aes(x = body_mass_g))+
  geom_histogram(bins = 10)+
  facet_wrap(~sex)

# QQ Plot
ggplot(chinstraps, aes(sample = body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)


# Descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarise(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

Get vectors of body_mass_g for female and male chinstrap penguins separately. 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)

```


### T-tests, a couple different ways

1. Is there a significant *difference* in body mass between male and female chinstrap penguins? (one-sided)

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```


2. Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins (two-sided)

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
#is male mass > female mass
```


Male chinstrap body mass is significanlty greater than female chinstrap body mass. 


## One-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```


Create a subset of only female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(penguin_f, aes(x=flipper_length_mm))+
  geom_histogram(aes(fill=species))+
  facet_wrap(~species, scales = "free") #not on same x,y scales

ggplot(penguin_f, aes(sample=flipper_length_mm))+
  geom_qq(aes(color=species))+
  facet_wrap(~species, scales = "free")
```


### Summary statistics

Starting from penguins_f. Create a summary table that groups by species and returns the following:

- mean of flipper length
- sd ""
- sample size
- variance 'var'
- standard error= sd/sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarise(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    n_flipper_length = n(),
    var_flipper_length = var(flipper_length_mm),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n())
  )

penguin_f_summary

leveneTest(flipper_length_mm ~ species, data = penguin_f)
```

Levene: null hypothesis that groups have equal variances
P = 0.06. Can retain null hypothesis.  Can assume variances are close enough. 

### Beeswarm plot with mean +/- 1sd

```{r}
ggplot()+
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5)+
  geom_point(data = penguin_f_summary, 
             aes(y = mean_flipper_length, x = species,
                 size=4))+
  geom_errorbar(data = penguin_f_summary, 
                aes(x = species, 
                    ymin = mean_flipper_length - sd_flipper_length, 
                    ymax = mean_flipper_length + sd_flipper_length),
                width = 0.1)+
  theme_light()
```



### One-way ANOVA
aov() runs type I anova (good if you have totally balanced factorial design); though type III anova better if you have unequal sample sizes and unbalanced design; often they end up being close.

Tukey's HSD. Does pairwise comparisons.

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```
Large F statistic value. Difference between groups very large compared to difference within groups. Would reject null hypoth that means for these groups are the same. 

P value sign for all groups.


### Type III ANOVA
Ex if have very unbalanced design (diff sample sizes within groups)

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```























