---
title: "Lab 5 - Penguin size comparisons"
author: "Jacey Van Wert"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(here)
library(effsize) #for cohen.d
library(broom)
library(car)
library(ggbeeswarm)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap using penguins dataset (preinstalled with {palmerpenguins})

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap") 


# Histograms
ggplot(chinstraps, aes(x = body_mass_g))+
  geom_histogram(bins = 10)+
  facet_wrap(~sex)

# QQ Plot
ggplot(chinstraps, aes(sample = body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)


# Descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarise(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

Get vectors of body_mass_g for female and male chinstrap penguins separately. 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)

```


### T-tests, a couple different ways

1. Is there a significant *difference* in body mass between male and female chinstrap penguins? (one-sided)

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```


2. Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins (two-sided)

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
#is male mass > female mass
```


Male chinstrap body mass is significanlty greater than female chinstrap body mass. 









